Amazonian Narratives in Presidential Discourse (1985 to 2020)

### 2021-11-08 11112021_Sposito_and_SilvaMuller_DRAFT.Rmd
#final formatting details # few more stop words

---
title: "Amazonian Narratives in Presidential Discourse (1985 to 2020)"
author: "Sposito & Silva-Muller"
date: "11th of November of 2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

@@

# 4 Preliminary Analysis

@@

library(dplyr)
library(stringi)
library(tm)
library(ggplot2)
library(tidytext)
library(kableExtra)
library(stringr)
library(wordcloud)
library(RColorBrewer)
library(tidyr)

@@

```{r speech frequencies, echo=FALSE, message=FALSE, warning=FALSE}

@@

ggplot(freq_year, aes(year, freqn))+
    geom_line()+
    geom_point()+
    geom_smooth(method=lm, color="#bab86c") +
      labs(x = "",
       y = "Relative share of speeches per year",
       title = "Presidential speeches mentioning the Amazon (1985-2020)",
       caption = "Frequencies normalized by the number of speeches per year in dataset",
       color = "")

freq_year <- dplyr::left_join(freq_year, amazon_def_year, by = "year") 
ggplot(freq_year, aes(x = year)) +
    geom_line(aes(y = freqn, color = "Frequency of amazonian speeches")) +
    geom_line(aes(y = `area km²`/100, color = "Annual deforestation (km²/100)")) +
      labs(x = "",
       y = "Relative share of speeches per year",
       title = "Presidential Speeches mentioning the Amazon (1985-2020)",
       subtitle = "Compared to annual deforestation rates since 1988",
       color = "")

```
From the first plot, the proportion of speeches that mention the word "Amazon" increase in frequency in 1989 (possibly related to Amazon Fires coverage from the New York Times see [here](https://www.nytimes.com/1988/08/12/world/vast-amazon-fires-man-made-linked-to-global-warming.html) and [here](https://www.nytimes.com/1989/03/23/world/brazil-smarting-from-the-outcry-over-the-amazon-charges-foreign-plot.html)), around 1992 (likely related to the Rio Conference), from 2005 to 2010 (possibly related to the "preservation" legislation turn under the Lula Administration), and after 2016 (possibly related to increasing deforestation rates and international attention). An interesting finding here is that while there is fluctuations along time, mentions to the Amazon did not increase or decrease, maintaining a general average of around 15% appearance in speeches. A second interesting finding is that mentions to the Amazon in presidential discourse do not correlate with deforestation rates. This might be due to different reasons: first, deforestation rate data is published once a year, so international response and presidential discourse are lagged. Relatedly, wider access to this data only started in the 2000s.A more interesting measure to use would be fire data or DETER data, which comes out monthly and might drive discourse. Second, deforestation data might correlate better with specific Amazonian narratives rather than mentions in general. We will test this idea in future versions.

@@

```{r top words by speaker, echo=FALSE, warning=FALSE, message=FALSE}

@@

# Let's see if bigrams help us more here
tamazon <- tidytext::tidy(amazon_corpus) %>%
  unnest_tokens(bigram, text , token = "ngrams", n = 2) %>%
  dplyr::count(bigram, sort = TRUE) %>%
  ungroup()
tamazon

# Let's get the 30 most frequent term by speaker
amazon_presid <- aggregate(Amazon_speeches$text, list(Amazon_speeches$president), paste, collapse = " ") %>%
  rename(doc_id = "Group.1", text = "x")
amazon_presid_t <- tm::VCorpus(DataframeSource(amazon_presid)) %>%
  tidy() %>% 
  unnest_tokens(word, text) %>%
  count(id, word, sort = TRUE)
Sarney <- head(dplyr::filter(amazon_presid_t, id == "Sarney"), 30) %>%
  mutate(Sarney = word) %>% select(Sarney)
Collor <- head(dplyr::filter(amazon_presid_t, id == "Collor"), 30) %>%
  mutate(Collor = word) %>% select(Collor)
Itamar <- head(dplyr::filter(amazon_presid_t, id == "Itamar"), 30) %>%
  mutate(Itamar = word) %>% select(Itamar)
FHC <- head(dplyr::filter(amazon_presid_t, id == "FHC"), 30) %>%  mutate(FHC = word) %>% select(FHC)
Lula <- head(dplyr::filter(amazon_presid_t, id == "Lula"), 30) %>%  mutate(Lula = word) %>% select(Lula)
Dilma <- head(dplyr::filter(amazon_presid_t, id == "Dilma"), 30) %>%  mutate(Dilma = word) %>% select(Dilma)
Temer <- head(dplyr::filter(amazon_presid_t, id == "Temer"), 30) %>%  mutate(Temer = word) %>% select(Temer)
Bolsonaro <- head(dplyr::filter(amazon_presid_t, id == "Bolsonaro"), 30) %>%
  mutate(Bolsonaro = word) %>% select(Bolsonaro)
f_speaker <- cbind(Sarney, Collor, Itamar, FHC, Lula, Dilma, Temer, Bolsonaro) %>% 
  kableExtra::kbl(caption = "Top 30 Words Per President") %>%
  kableExtra::kable_classic(full_width = F, html_font = "Times New Roman")
f_speaker
```

@@

@@

sovereignty <- ("amazonia e brasileira|amazonia e nossa|soberania|interesse estrangeiro|ocupar|forcas armadas|militar|fronteir|patrimonio|nacional|defesa|exercito|dono|nossa")
development <- ("estrada|rodovia|hidroeletrica|desenvolv|balbina|itaipu|incentivos fiscais|integrar|integracao|zona franca|transamazonica|madeira|porto|madeireira|produzir")
conservation <- ("preserv|conserv|determinacao|povos indigenas|indigenas|direitos humanos|areas demarcadas|queimada|sustentav|meio ambiente|florest|cooperaçao|protecao|proteger")
anti_environmentalism <- ("preservamos demais|nao vamos demarcar|demarcar menos|agricultura|producao de gado|ambientalistas|ong|xiita") # This appears harder to operationalize than others...

@@

@@

ggplot(am_narratives, aes(x = year, y = value , fill = narrative)) +
  geom_line(aes(group = narrative, color = narrative), size = 1) +
  labs(x = "Year",
       y = "",
       title = "Narratives in Presidential Speeches per year since 1985 in Brazil")+
    facet_wrap(~narrative)+
    theme (legend.position="none")

@@

Lahsen, Myanna. 2009. “A Science–Policy Interface in the Global South: The Politics of Carbon Sinks and Science in Brazil.” Climatic Change 97 (3–4): 339–72.                                    https://doi.org/10.1007/s10584-009-9610-6.

@@

>>>>

---
title: "Amazonian Narratives in Presidential Discourse (1985 to 2020)"
author:
- Henrique Sposito^[Phd Candidate, The Graduate Institute Geneva, henrique.sposito@graduateinstitute.ch]
- Livio Silva-Muller^[Phd Candidate, The Graduate Institute Geneva, livio.silva@graduateinstitute.ch]
date: "11th of November of 2021"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---

@@

# 4 Exploratory analysis

In the following, we will show our work in progress in terms of the empirical analysis. At the moment, we are trying to figure out what methods we will actually use and why. The number of speeches mentioning the Amazon (n=861), is not high enough to justify certain machine learning approaches as manually coding a fraction. We tested more inductive approaches (topic modelling), and it seems to work well, but at the same time the text volume is too small to justify. As there are other factors we are interested at (e.g. a measure of intensity, how (a)permissive and  (b)explicit is the discourse), we think a combination of manual coding and more quantitative approach might help.

@@

library(dplyr)
library(stringi)
library(tm)
library(ggplot2)
library(tidytext)
library(kableExtra)
library(stringr)
library(wordcloud)
library(RColorBrewer)
library(tidyr)
library(formatR)

@@

```{r speech frequencies I, echo=FALSE, message=FALSE, warning=FALSE}

@@

ggplot(freq_year, aes(year, freqn))+
    geom_line()+
    geom_point()+
    geom_smooth(method=lm, color="#bab86c") +
      labs(x = "",
       y = "Relative share of speeches per year",
       title = "Presidential speeches mentioning the Amazon (1985-2020)",
       caption = "Frequencies normalized by the number of speeches per year in dataset",
       color = "")
```

```{r speech frequencies II, echo=FALSE, message=FALSE, warning=FALSE}

freq_year <- dplyr::left_join(freq_year, amazon_def_year, by = "year") 
ggplot(freq_year, aes(x = year)) +
    geom_line(aes(y = freqn, color = "Frequency of amazonian speeches")) +
    geom_line(aes(y = `area km²`/100, color = "Annual deforestation (km²/100)")) +
      labs(x = "",
       y = "Relative share of speeches per year",
       title = "Presidential Speeches mentioning the Amazon (1985-2020)",
       subtitle = "Compared to annual deforestation rates since 1988",
       color = "")+
       theme(legend.position ="bottom")

From the first plot, the proportion of speeches that mention the word "Amazon" increase in frequency in 1989 (possibly related to Amazon Fires coverage from the New York Times see [here](https://www.nytimes.com/1988/08/12/world/vast-amazon-fires-man-made-linked-to-global-warming.html) and [here](https://www.nytimes.com/1989/03/23/world/brazil-smarting-from-the-outcry-over-the-amazon-charges-foreign-plot.html)), around 1992 (likely related to the Rio Conference), from 2005 to 2010 (possibly related to the "preservation" legislation turn under the Lula Administration), and after 2016 (possibly related to increasing deforestation rates and international attention). An interesting finding here is that while there is fluctuations along time, mentions to the Amazon did not increase or decrease, maintaining a general average of around 15% appearance in speeches. A second interesting finding is that mentions to the Amazon in presidential discourse do not correlate with deforestation rates. This might be due to different reasons: first, deforestation rate data is published once a year, so international response and presidential discourse are lagged. Relatedly, wider access to this data only started in the 2000s.A more interesting measure to use would be fire data or DETER data, which comes out monthly and might drive discourse. Second, deforestation data might correlate better with specific Amazonian narratives rather than mentions in general. We will test this idea in future versions.

@@

```{r top words by speaker I, echo=FALSE, warning=FALSE, message=FALSE, fig.height = 6}

@@

# Let's see if bigrams help us more here
tamazon <- tidytext::tidy(amazon_corpus) %>%
  unnest_tokens(bigram, text , token = "ngrams", n = 2) %>%
  dplyr::count(bigram, sort = TRUE) %>%
  ungroup()
tamazon
```

In terms of most frequent words, we already observe mentions related to the narratives we expect (develop, integrate, produce, deforestation).  Pairs of words (bigrams) also point towards the same (meio ambiente, forcas armadas, zona franca, desenvolvimento sustentavel).

```{r top words by speaker II, echo=FALSE, warning=FALSE, message=FALSE}

# Let's get the 30 most frequent term by speaker
amazon_presid <- aggregate(Amazon_speeches$text, list(Amazon_speeches$president), paste, collapse = " ") %>%
  rename(doc_id = "Group.1", text = "x")
amazon_presid_t <- tm::VCorpus(DataframeSource(amazon_presid)) %>%
  tidy() %>% 
  unnest_tokens(word, text) %>%
  count(id, word, sort = TRUE)
Sarney <- head(dplyr::filter(amazon_presid_t, id == "Sarney"), 30) %>%
  mutate(Sarney = word) %>% select(Sarney)
Collor <- head(dplyr::filter(amazon_presid_t, id == "Collor"), 30) %>%
  mutate(Collor = word) %>% select(Collor)
Itamar <- head(dplyr::filter(amazon_presid_t, id == "Itamar"), 30) %>%
  mutate(Itamar = word) %>% select(Itamar)
FHC <- head(dplyr::filter(amazon_presid_t, id == "FHC"), 30) %>%  mutate(FHC = word) %>% select(FHC)
Lula <- head(dplyr::filter(amazon_presid_t, id == "Lula"), 30) %>%  mutate(Lula = word) %>% select(Lula)
Dilma <- head(dplyr::filter(amazon_presid_t, id == "Dilma"), 30) %>%  mutate(Dilma = word) %>% select(Dilma)
Temer <- head(dplyr::filter(amazon_presid_t, id == "Temer"), 30) %>%  mutate(Temer = word) %>% select(Temer)
Bolsonaro <- head(dplyr::filter(amazon_presid_t, id == "Bolsonaro"), 30) %>%
  mutate(Bolsonaro = word) %>% select(Bolsonaro)
f_speaker <- cbind(Sarney, Collor, Itamar, FHC, Lula, Dilma, Temer, Bolsonaro) %>% 
  kableExtra::kbl(caption = "Top 30 Words Per President") %>%
  kableExtra::kable_classic(full_width = F, html_font = "Times New Roman")

landscape(f_speaker, margin=NULL)
```

Moving to most mentioned words by presidents, there are some interesting anectodal findings. For instance, deforestation appears to be more salient in the vocabulary employed by Lula, Dilma and Temer versus their counterparts. The words "soberania" and "military" only appears as part of Bolsonaro's vocabulary.Desenvolvimento, on the other hand, seems to be part of all presidents vocabulary except of Temer.

@@

The tabel 2 (above) and table 3(below) portray the result of a structural topic model. We utilize LDA (Latent Dirichlet allocation), which is a model for estimating the mixture of words associated with a topic (cluster) and determining the probabilities of topics in each mention, we start to see some interesting patterns (for more information on LDA and R approaches [here](https://www.tidytextmining.com/index.html)). We set the number of clusters to be found (topics) at 10, randomly. We, then, select the top ten words coded for each topic.

Development appears in 6 out of 10 topics, suggesting it is a cross-cutting vocabulary. In its purest form, we observe the developmentalist narrative in topic 1 and 3. Three topics relate to themes of security, development, and integration (4,7,8) and one other touches on sovereignty in its purest form (9), which combined understand as the sovereignty narrative. Two topics portray vocabulary that we understand as constituting the conservation narrative (2 and 10), with topic 10 being the purest form in the text.

Once the clusters are extracted, we can indicate, probabilistically, how presidents' discourse fall within a topic, that's what the second table does for the 2 most salient topics per speakier.We see that, for example, Lula's discourses about the Amazon often is clustered with themes of development and of deforestation. At the same time, Bolsonaro's discourses about the Amazon often is clustered with themes of integration, energy, and the military forces. Sarney engaged with themes of preservation, while Cardoso engages with themes related to energy and integration. Others as Rousseff and Temer appear to engage with themes of geography, politics, energy; while Collor an Itamar appear to engage with themes of security, development, and integration.

@@

sovereignty <- ("amazonia e brasileira|amazonia e nossa|soberania|interesse estrangeiro|
                ocupar|forcas armadas|militar|fronteir|patrimonio|nacional|defesa|exercito|
                dono|nossa")

development <- ("estrada|rodovia|hidroeletrica|desenvolv|balbina|itaipu|
                incentivos fiscais| integrar|integracao|zona franca|transamazonica|madeira|
                porto|madeireira|produz|energia")

conservation <- ("preserv|conserv|determinacao|povos indigenas|indigenas|direitos humanos|
                 areas demarcadas|queimada|sustentav|meio ambiente|florest|cooperaçao|
                 protecao|proteg")

anti_environmentalism <- ("preservamos demais|nao vamos demarcar|demarcar menos|
                          agricultura|producao de gado|ambientalistas|ong|xiita")

@@

Shifting our methodological strategy from an inductive approach to a more deductive and theoretically driven analysis of the text, leads similar results.We operationalize our narratives by deductively creating four groups of words (dictionaries) and assigning each contextualized mention to the Amazon to a narrative. The dictionaries are presented in the code chunk above. Our narratives approach, while still empirically weak, already shows some interesting trends (table 4, table 5, and narratives plots; below). First, the 1988 NYC pictures of the Amazon burning piked all narratives during the Sarney presidency. Second, while Lula was the president that most performed conservation narratives until 2010, Bolsonaro mobilized sovereignty narratives strongly. Third, generally the developmentalist narrative have decreased along time but has appeared most frequently during the Lula mandate, suggesting sustainable development was at its highest in those years. Fourth, while the anti socio-environmentalist narrative was generally not so frequent (this might be an identification problem or that it is simply a sub-type of the developmentalist narrative), it was present across all speakers with Bolsonaro using it the most.

@@

ggplot(am_narratives, aes(x = year, y = value , fill = narrative)) +
  geom_line(aes(group = narrative, color = narrative), size = 1) +
  labs(x = "Year",
       y = "",
       title = "Narratives in Presidential Speeches per year since 1985 in Brazil")+
    geom_smooth(method=lm, se=FALSE, color="black", linetype="dashed",size=0.5)+
    facet_wrap(~narrative)+
    theme (legend.position="none")

@@

Lahsen, Myanna. (2009). “A Science–Policy Interface in the Global South: The Politics of Carbon Sinks and Science in Brazil.” Climatic Change 97 (3–4): 339–72.                                    https://doi.org/10.1007/s10584-009-9610-6.

@@

Silva-Muller, L., & Faul, M. (2021) “Protecting the Amazon and its people: the role of civil society in the local effectiveness of transnational partnerships”. In Partnerships for                        Sustainability: Pathways to Effectiveness. Edt. by Andonova, L., Faul, M. and Piselli, D. Routledge: London.

Silva-Muller, L. (forthcoming) "Payment for ecosystem services and the practices of environmental fieldworkers in policy-implementation: the case of Bolsa Floresta in the Brazilian                        Amazon"
